---
title: "Data Understanding"
format: html
editor: visual
---

Adding to the foundation of Business Understanding, the Data Understanding phase focuses on identifying, collecting, and analyzing data sets that can help the project. This phase also has four tasks:

See table with variables and further information!

## Collect initial data: Acquire the necessary data and (if necessary) load it into your analysis tool.

**Data overview**

```{r}
# Load  dataset with  separator ";", a header and empty cells recognized as NAs
data <- read.csv("marketing.csv", sep = ";", header = TRUE, na.strings = "")

# Check the data set------------------------------------------------------------
head(data) # Displays the first lines of the data set
str(data) # Outputs a summary of the data structure
nrow(data) # cross checking number or rows
```

-\>17 variables (columns) and 41’188 observations (rows)

-\> Types: chr, int, and num

**Missing values (NAs)**

***Variable Balance is not in the dataset !***

```{r}
# Determine the number of missing values for each variable----------------------
missing_values <- sapply(data, function(x) sum(is.na(x)))
missing_values # Show the number of missing values for each variable
```

**Relationships between variables:** previous = 0 & poutcome = nonexistent

```{r}
# Relationship between previous = 0 & poutcome = nonexistent--------------------
# Number of previous = 0
count_previous_0 <- sum(data$previous == 0) 
count_previous_0

# Number of poutcome ="nonexistent"
count_poutcome_nonexistent <- sum(data$poutcome == "nonexistent") 
count_poutcome_nonexistent

# Number of previous = 0 & poutcome = "nonexistent" at the same time
count <- sum(data$poutcome[data$previous == 0] == "nonexistent") 
count

# check if relationship is true (all numbers are equal)
realtionship <- (count_previous_0 == count_poutcome_nonexistent) && +
  (count_previous_0 == count) && (count_poutcome_nonexistent == count)
realtionship
```

1.  Describe data: Examine the data and document its surface properties like data format, number of records, or field identities.

    See table with variables and further information!

    **Statistical Analysis**

    **of numeric variables**

    ```{r}
    # statistical analysis of the data----------------------------------------------
    # statistical Summary of numeric variables
    summary_numeric <- summary(data[, sapply(data, is.numeric)]) 
    summary_numeric

    # standard deviation of of numeric variables
    sd_numeric <- sapply(data[, sapply(data, is.numeric)], sd)
    sd_numeric
    ```

    **of discrete** **numeric variables (integers)**

    ```{r}
    # Frequency of the individual values for discrete numeric variables (integers)
    frequency_int_data <- lapply(data[, sapply(data, is.integer)], table)
    frequency_int_data

    # Unique values for discrete numeric variables (integers)
    unique_int_values <- sapply(data[, sapply(data, is.integer)], unique)
    unique_int_values
    ```

    **of non-numeric variables**

    ```{r}
    # Frequency of individual categories for non-numeric variables including NA
    frequency_data <- lapply(data[, sapply(data, is.character)], function(x) table(x, useNA = "always"))
    frequency_data

    # Number of unique values for non-numeric variables
    unique_data <- sapply(data[, sapply(data, is.character)], function(x) length(unique(x)))
    unique_data
    ```

2.  Explore data: Dig deeper into the data. Query it, visualize it, and identify relationships among the data.

3.  Verify data quality: How clean/dirty is the data? Document any quality issues.
