---
title: "Data Understanding"
format: docx
prefer-html: true
toc: true
toc-depth: 5
editor: visual
warning: false
---

Adding to the foundation of Business Understanding, the Data Understanding phase focuses on identifying, collecting, and analyzing data sets that can help the project. This phase also has four tasks:

See *Variables Overview.xlsx* for further information!

```{r}
# load libraries, that might be needed
library(knitr)      # web widget
library(tidyverse)  # data manipulation
library(data.table) # fast file reading
library(caret)      # rocr analysis
library(ROCR)       # rocr analysis
library(kableExtra) # nice table html formating 
library(gridExtra)  # arranging ggplot in grid
library(rpart)      # decision tree
library(rpart.plot) # decision tree plotting
library(caTools)    # split 
```

## 1. Collect initial data: Acquire the necessary data and (if necessary) load it into your analysis tool.

### **Data overview**

```{r}
# Load  dataset with  separator ";", a header and empty cells recognized as NAs
data <- read.csv("marketing.csv", sep = ";", header = TRUE, na.strings = "")

# Check the data set
names(data) # Displays names of attributes
head(data) # Displays the first lines of the data set
str(data) # Outputs a summary of the data structure
nrow(data) # cross checking number or rows
dim(data) # cross checking the dimension
```

### Missing Variables

Variable Balance is not in the dataset !

### **Check for Duplicate Rows**

```{r}
# Check for Duplicate Rows
sum(duplicated(data))
```

### **Missing values (NAs)**

```{r}
# How many rows have missing data
sum(!complete.cases(data))

# How Many Rows Are Completely Missing Values In All Columns
all.empty = rowSums(is.na(data))==ncol(data)
sum(all.empty)

# Determine the number of missing values for each variable
missing_values <- sapply(data, function(x) sum(is.na(x)))
missing_values # Show the number of missing values for each variable
```

## 2 Describe data: Examine the data and document its surface properties like data format, number of records, or field identities.

See *Variables Overview.xlsx* for further information!

### **Statistical Analysis**

#### **of numeric variables**

```{r}
# statistical analysis of the data
# statistical Summary of numeric variables
summary_numeric <- summary(data[, sapply(data, is.numeric)]) 
summary_numeric

# standard deviation of of numeric variables
sd_numeric <- sapply(data[, sapply(data, is.numeric)], sd)
sd_numeric
```

#### **of discrete** **numeric variables (integers)**

```{r}
# Frequency of the individual values for discrete numeric variables (integers)
frequency_int_data <- lapply(data[, sapply(data, is.integer)], table)
frequency_int_data

# Unique values for discrete numeric variables (integers)
unique_int_values <- sapply(data[, sapply(data, is.integer)], unique)
unique_int_values
```

#### **of non-numeric variables**

```{r}
# Frequency of individual categories for non-numeric variables including NA
frequency_data <- lapply(data[, sapply(data, is.character)], function(x) table(x, useNA = "always"))
frequency_data

# Number of unique values for non-numeric variables
unique_data <- sapply(data[, sapply(data, is.character)], function(x) length(unique(x)))
unique_data
```

## 3. Explore data: Dig deeper into the data. Query it, visualize it, and identify relationships among the data.

### **Visualizations**

**Histograms**: Visualize the distribution of numeric variables.

-   **Density:** Density scales the height of the bars based on the relative frequency of data within each bin. It's useful for comparing the shape of distributions across different datasets or variables, independent of the number of observations or bin width.

-   **Frequency:** Frequency displays the absolute number of observations in each bin. It's useful for understanding the exact count of observations in specific areas of the distribution, particularly for identifying outliers or regions with unusually high or low frequencies.

#### **AGE**

```{r}
# stattistical summary of Age
summary(data$age)

# Histogram count of Ages
gg = ggplot (data) 
p1 = gg + geom_histogram(aes(x=age),color="black", fill="white", binwidth = 5) +
  ggtitle('Age Distribution (red mean line)') +
  ylab('Count') +
  xlab('Age') +
  geom_vline(aes(xintercept = mean(age), color = "red")) +
  scale_x_continuous(breaks = seq(0,100,5)) +
  theme(legend.position = "none")
p1
```

```{r}
# Histogram y = Frequency, x = Age
hist(data$age, main = "Histogram of Age", xlab = "Age", breaks = unique(data$age), freq = TRUE)

# Histogram y = Density, x = Age
hist(data$age, main = "Histogram of Age", xlab = "Age", breaks = unique(data$age), freq = FALSE)

# Boxplot for Age
boxplot(data$age, horizontal = TRUE, main = "Boxplot of Age", xlab = "Age")
```

#### **JOB**

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_job <- c('admin.', 'blue-collar', 'entrepreneur', 'housemaid', 'management', 'retired', 'self-employed', 'services', 'student', 'technician', 'unemployed', 'unknown')

# Barplot for Job Categories
barplot(table(data$job, useNA = "always"), main = "Barplot of Job Categories", xlab = "Job Category", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_job, mgp = c(3, 1, 0))
```

#### **MARITAL**

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_marital <- c('divorced', 'married', 'single', 'unknown')

# Barplot for Marital Status
barplot(table(data$marital, useNA = "always"), main = "Barplot of Marital Status", xlab = "Marital Status", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_marital, mgp = c(3, 1, 0))

```

#### **EDUCATION**

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_education <- c('illiterate', 'basic.4y', 'basic.y', 'basic.9y', 'high.school', 'professional.course', 'university.degree', 'unknown')

# Barplot for Education Level
barplot(table(data$education, useNA = "always"), main = "Barplot of Education Level", xlab = "Education Level", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_education, mgp = c(3, 1, 0))
```

#### **DEFAULT**

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_default <- c('yes', 'no', 'unknown')

# Barplot for Credit in Default
barplot(table(data$default, useNA = "always"), main = "Barplot of Credit in Default", xlab = "Credit in Default", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_default, mgp = c(3, 1, 0))
```

#### **HOUSING**

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_housing <- c('yes', 'no', 'unknown')

# Barplot for Housing Loans
barplot(table(data$housing, useNA = "always"), main = "Barplot of Housing Loans", xlab = "Housing Loans", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_housing, mgp = c(3, 1, 0))
```

#### **LOAN**

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_loan <- c('yes', 'no', 'unknown')

# Barplot for Personal Loans
barplot(table(data$loan, useNA = "always"), main = "Barplot of Personal Loans", xlab = "Personal Loans", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_loan, mgp = c(3, 1, 0))
```

#### **DAY OF WEEK**

```{r}
# Define Categories (for a nice order)
categories_days <- c('mon', 'tue', 'wed', 'thu', 'fri')


# Barplot for Day of Week
barplot(table(data$day_of_week), main = "Barplot of Days of Week", xlab = "Days of Week", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_days, mgp = c(3, 1, 0))
```

#### **MONTH**

```{r}
# Define Categories (for a nice order)
categories_months <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

# Barplot for Month
barplot(table(data$month), main = "Barplot of Months", xlab = "Months", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_months, mgp = c(3, 1, 0))
```

#### **CAMPAIN**

```{r}
# Histogram y = Frequency, x = Campain
hist(data$campaign, main = "Histogram of Number of Contacts during this Campain", xlab = "Number of Contacts", breaks = unique(data$campaign), freq = TRUE)

# Histogram y = Density, x = Campain
hist(data$campaign, main = "Histogram of Number of Contacts during this Campain", xlab = "Number of Contacts", breaks = unique(data$campaign), freq = FALSE)

# Boxplot for Campain
boxplot(data$campaign, horizontal = TRUE, main = "Boxplot of Number of Contacts during this Campain", xlab = "Number of Contacts")
```

#### **PREVIOUS**

```{r}
# Histogram y = Frequency, x = Previous
hist(data$previous, main = "Histogram of Number of Contacts before this Campain", xlab = "Number of Previous Contacts", breaks = unique(data$previous), freq = TRUE)

# Histogram y = Density, x = Previous
hist(data$previous, main = "Histogram of Number of Contacts before this Campain", xlab = "Number of Previous Contacts", breaks = unique(data$previous), freq = FALSE)

# Boxplot for Previous
boxplot(data$previous, horizontal = TRUE, main = "Boxplot of Number of Contacts before this Campain", xlab = "Number of Previous Contacts")
```

#### **POUTCOME**

```{r}
# Define Categories (for a nice order)
categories_poutcome <- c("success", "failure", "nonexistent")

# Barplot for Poutcome
barplot(table(data$poutcome), main = "Barplot of Outcome of the previous marketing campaign", xlab = "Outcome", ylab = "Frequency", las = 2, cex.names = 0.75, names.arg = categories_poutcome, mgp = c(3, 1, 0))
```

#### **Emp.var.rate =** Employment variation rate – quarterly indicator

```{r}
# Histogram y = Frequency, x = Emp.var.rate 
hist(data$emp.var.rate, main = "Histogram of Employment variation rate", xlab = "Rate", breaks = unique(data$emp.var.rate), freq = TRUE)

# Histogram y = Density, x = Emp.var.rate 
hist(data$emp.var.rate, main = "Histogram of Employment variation rate", xlab = "Rate", breaks = unique(data$emp.var.rate), freq = FALSE)

# Boxplot for Emp.var.rate 
boxplot(data$emp.var.rate, horizontal = TRUE, main = "Boxplot of Employment variation rate", xlab = "Rate")
```

#### **Cons.price.idx =** Consumer price index – monthly indicator

```{r}
# Histogram y = Frequency, x = Cons.price.idx 
hist(data$cons.price.idx, main = "Histogram of Consumer price index", xlab = "Index", breaks = unique(data$cons.price.idx), freq = TRUE)

# Histogram y = Density, x = Cons.price.idx 
hist(data$cons.price.idx, main = "Histogram of Consumer price index", xlab = "Index", breaks = unique(data$cons.price.idx), freq = FALSE)

# Boxplot for Cons.price.idx
boxplot(data$cons.price.idx, horizontal = TRUE, main = "Boxplot of Consumer price index", xlab = "Index")
```

#### **Cons.conf.idx =** Consumer confidence index – monthly indicator

```{r}
# Histogram y = Frequency, x = Cons.conf.idx 
hist(data$cons.conf.idx, main = "Histogram of Consumer confidence index", xlab = "Index", breaks = unique(data$cons.conf.idx), freq = TRUE)

# Histogram y = Density, x = Cons.conf.idx 
hist(data$cons.conf.idx, main = "Histogram of Consumer confidence index", xlab = "Index", breaks = unique(data$cons.conf.idx), freq = FALSE)

# Boxplot for Cons.price.idx
boxplot(data$cons.conf.idx, horizontal = TRUE, main = "Boxplot of Consumer confidence index", xlab = "Index")
```

#### **Euribor3m =** Euribor 3-month rate – daily indicator

```{r}
# Histogram y = Frequency, x = Euribor3m  
hist(data$euribor3m, main = "Histogram of Euribor 3-month rate", xlab = "Rate", breaks = unique(data$euribor3m), freq = TRUE)

# Histogram y = Density, x = Euribor3m  
hist(data$euribor3m, main = "Histogram of Euribor 3-month rate", xlab = "Rate", breaks = unique(data$euribor3m), freq = FALSE)

# Boxplot for Euribor3m 
boxplot(data$euribor3m, horizontal = TRUE, main = "Boxplot of Euribor 3-month rate", xlab = "Rate")
```

#### **Y** - Target Variable: Has the client subscribed a term deposit

```{r}
# Barplot for Y
barplot(table(data$y), main = "Barplot of Subscription made", xlab = "Subscription made", ylab = "Frequency", las = 2, cex.names = 0.75, mgp = c(3, 1, 0))
```

##### **Outcome Imbalance**

Observe that the dataset predicted outcome (y) is skewed towards ‘no’ with over 88.7%.

```{r}
prop.table(table(data$y))
```

### **Relationships between variables:**

#### previous & poutcome

```{r}
# Reorder levels of poutcome
data$poutcome <- factor(data$poutcome, levels = c("nonexistent", "success", "failure"))

# Create the scatter plot
ggplot(data, aes(x = poutcome, y = previous)) +
  geom_jitter(width = 0.3, height = 0) +  # Add jitter to prevent overlapping
  labs(title = "Scatter Plot of Poutcome and Previous", x = "Poutcome", y = "Previous Contacts")
```

##### previous = 0 & poutcome = nonexistent

```{r}
# Relationship between previous = 0 & poutcome = nonexistent
# Number of previous = 0
count_previous_0 <- sum(data$previous == 0) 
count_previous_0

# Number of poutcome ="nonexistent"
count_poutcome_nonexistent <- sum(data$poutcome == "nonexistent") 
count_poutcome_nonexistent

# Number of previous = 0 & poutcome = "nonexistent" at the same time
count <- sum(data$poutcome[data$previous == 0] == "nonexistent") 
count

# check if relationship is true (all numbers are equal)
realtionship <- (count_previous_0 == count_poutcome_nonexistent) && +
  (count_previous_0 == count) && (count_poutcome_nonexistent == count)
realtionship
```

### **Age Distribution vs Marital Status That Subscribes Term Deposit**

*The bulk of clients are married or divorced.* Sharp drop of clients above age 60 with marital status ‘divorced’ and ‘married’. \*Single clients drop in numbers above age 30-40.

```{r}
ggplot(data, aes(x=age, fill=marital)) + 
  geom_histogram(binwidth = 2, alpha=0.7, na.rm = TRUE) +
  facet_grid(cols = vars(y)) +
  expand_limits(x=c(0,100)) +
  scale_x_continuous(breaks = seq(0,100,10)) +
  ggtitle("Age Distribution by Marital Status")
```

### **Age vs Subscription**

Most clients that subscribe are between age 25 to 55. Mean age for all clients is at age 40 years.

```{r}
mean(data$age)

mu <- data %>% group_by(y) %>% summarize(grp.mean=mean(age))

ggplot (data, aes(x=age)) + 
  geom_histogram(color = "blue", fill = "blue", binwidth = 5) +
  facet_grid(cols=vars(y)) + 
  ggtitle('Age Distribution by Subscription (red line is mean') + ylab('Count') + xlab('Age') +
  scale_x_continuous(breaks = seq(0,100,5)) +
  geom_vline(data=mu, aes(xintercept=grp.mean), color="red", linetype="dashed")
```

### **Education vs Subscription**

Having higher education is seen to contribute to higher subscription of term deposit. Most clients who subscribe are from ‘high school’ and ‘university degree’ education levels.

```{r}
# Define Categories (so emtpy gets shown as unknown)
categories_education <- c('illiterate', 'basic.4y', 'basic.y', 'basic.9y', 'high.school', 'professional.course', 'university.degree', 'unknown')

ggplot(data = data, aes(x=factor(education, categories_education), fill=y)) +
         geom_bar(na.rm = TRUE) +
         ggtitle("Term Deposit Subscription based on Education Level") +
         xlab(" Education Level") +
         guides(fill=guide_legend(title="Subscription of Term Deposit")) +
         theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

The following table presents the number of subscriptions per education level. In total, 4640 clients have subscribed. The 'percentage' column showcases the proportion of subscribed clients relative to the total number of subscriptions, expressed as a percentage. Notably, individuals with higher education levels exhibit higher subscription rates: high school (22.22%), professional course (12.82%), and university degree (35.99%).

```{r}
# categories to have the correct order
categories_education <- c('illiterate', 'basic.4y', 'basic.6y', 'basic.9y', 'high.school', 'professional.course', 'university.degree', 'unknown')

# Filtering the data to obtain only customers with a subscription
subscribed_customers <- data[data$y == "yes", ]

# Total number of subscribed customers
total_subscribed <- nrow(subscribed_customers)

# Grouping data by education level & counting subscribed customers in each category
education_counts <- subscribed_customers %>%
  group_by(education) %>%
  summarise(subscribed_customers = n())

# Calculating percentage of subscribed customers for each education level
education_counts <- education_counts %>%
  mutate(percentage = round(subscribed_customers / total_subscribed * 100, 2))

# Sorting data according to the order of categories
education_counts <- education_counts[order(match(education_counts$education, categories_education)), ]

# Printing results
print(education_counts)
```

The table below displays the number of subscriptions and total clients across different education levels. The 'percentage' column indicates the proportion of subscribed clients per education level relative to the total number of clients per education level, presented as a percentage. Notably, individuals with higher education levels exhibit higher subscription rates, such as those with a high school education (10.84%), professional course (11.35%), and university degree (13.72%).

However, it's essential to consider the context. For instance, the illiterate group shows a high subscription percentage (22.22%), but this is based on a small sample size of only 18 out of 41188 clients contacted. Similarly, the 14.50% subscription rate among clients with an unknown education level appears relatively high, yet this is based on a relatively small group of 1731 clients with an unknown education level.

```{r}
# Define order of education categories
categories_education <- c('illiterate', 'basic.4y', 'basic.6y', 'basic.9y', 'high.school', 'professional.course', 'university.degree', 'unknown')

# Filter data to obtain only customers with successful subscriptions
subscribed_customers <- data[data$y == "yes", ]

# Calculate total number of subscribed customers
total_subscribed <- nrow(subscribed_customers)

# Group data by education level & count subscribed customers in each category
subscription_counts <- subscribed_customers %>%
  group_by(education) %>%
  summarise(subscribed = n())

# Calculate total number of customers per education level
total_clients <- data %>%
  group_by(education) %>%
  summarise(total = n())

# Merge the two datasets
subscription_summary <- merge(subscription_counts, total_clients, by = "education", all = TRUE)

# Calculate percentage of subscriptions relative to total number of customers
subscription_summary$percentage <- round(with(subscription_summary, subscribed / total * 100), 2)

# Adjust order of rows based on predefined education categories
subscription_summary <- subscription_summary %>%
  arrange(match(education, categories_education))

# Print results
print(subscription_summary)
```

### **Subscription based on Number of Contact during Campaign**

It can be observed from barchart that there will be nearly no subscription beyond 10 contact during the campaign. Future campaign could improve resource utilization by setting limits to contacts during a campaign. Future campaigns can focus on first 3 contacts as it will have higher subscription rate.

```{r}
ggplot(data=data, aes(x=campaign, fill=y))+
  geom_histogram(na.rm = TRUE, bins = 30)+
  ggtitle("Subscription based on Number of Contacts during the Campaign")+
  xlab("Number of Contacts during the Campaign")+
  xlim(c(min=1,max=25)) +
  guides(fill=guide_legend(title="Subscription of Term Deposit"))

```

```{r}
# Grouping data by campaign number and counting customers in each category
campaign_counts <- data %>%
  group_by(campaign) %>%
  summarise(count_campaign = n(),
            count_y_yes = sum(y == "yes"),
            count_y_no = sum(y == "no"))

# Output results
print(campaign_counts)

```

xxxxxx hier weiter machen

## 4. Verify data quality: How clean/dirty is the data? Document any quality issues.

17 variables (columns) and 41’188 observations (rows)

There is some missing variables (balance), and values, also wome rows are duplicated.

-\> missing variable balance: is that on purpose?

-\> missing values: some/all should be filled with default value (example mean).

-\> delete duplicated rows.

Types: chr, int, and num

-\> some types might need to be changed e.g. as.factor(), ex. : housing : Factor w/ 2 levels "no","yes": 1 2
